\boldmath Primordial Black Hole Formation in a Dust Bouncing Model

### 2024-09-30 referee_response.tex

\documentclass[a4paper,11pt]{article}
\pdfoutput=1 % if your are submitting a pdflatex (i.e. if you have
             % images in pdf, png or jpg format)

\usepackage[T1]{fontenc} % if needed

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% My Packages %%%%%%%%%%%%%%%%%
\usepackage[english]{babel}
\usepackage{braket}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{physics,slashed}
\usepackage{graphicx}
\usepackage[toc,page]{appendix}
\usepackage{hyperref}
\usepackage[noabbrev]{cleveref}

\usepackage{xcolor}
\usepackage{mdframed}

\usepackage[hmarginratio=1:1]{geometry}

\begin{document}
We appreciate the interest in our paper and thank you for the time taken to review it. Regarding the points suggested, here are our thoughts on the matter:

1 - a) To obtain the matter Lagrangian for this model, we employ the Schutz formalism. It assumes that the fluid's velocity can be expressed in terms of potentials of scalar fields, and the fluid's enthalpy is determined by the magnitude of the velocity. This allows us to write the Lagrangian as a functional of these fields. A detailed application of this formalism can be found in Ref.~\cite{fluidgeral}. We have included a brief description of this approach in Section 2.2 and corrected the cited references.

1 - b) Indeed, if the sound speed were to approach zero, we would encounter a divergence problem. This is why we limit ourselves to small but non-zero equations of state for dust, which is in agreement with the literature. Regarding the $\zeta$ field, we can observe from the initial conditions in equation 3.32, along with the definitions of 
z and the frequency, that the modes for $\zeta$ will evolve proportionally to 
$\frac{1}{\sqrt{m \omega_k}} \propto c_s$, meaning that higher-order perturbative terms would evolve similarly.

In \cite{vitenti}, the authors demonstrate that the first-order terms in the perturbative expansion are indeed smaller than the zero-order terms. To replicate this for higher-order terms, we would need to prove that all third-order terms are smaller than second-order terms in a general manner. Simply comparing one term in the expansion, such as $\zeta^3$ to $\zeta^2$, does not imply that all second-order terms in the Lagrangian are less significant than third-order terms. This kind of proof has not yet been addressed in the literature and is beyond the scope of this paper.

We have reorganized Sections 4.3 and 4.4 and considered a very small, but non-zero, equation of state for dark matter to analyze the pressureless case.

2 - a) DEMETRIO ADICIONAR

2 - b) SANDRO ADICIONAR In our quantum model, we observe that there is a quantum contribution to the usual Friedmann equation. This contribution should not be interpreted as a new fluid but rather as an effective correction. Essentially, we use the classical fluid density with the scale factor following the Bohm trajectory.

Moreover, if we compute the time in the universe when the quantum contribution reaches 10\% of the classical contribution, we find that this occurs only at scales very close to the bounce. At these timescales, the primordial black holes formed would have such small masses that they would evaporate quickly and thus would not affect our analysis. We restrict our study to times well before the bounce, where such quantum corrections are not significant. Therefore, we only consider $w_{matter}$ and not $w_{eff}$.

We have added the calculation of the times when the quantum contribution becomes 10\% of the classical one to the paper.

2 - c) At bouncing point, we have that \bar{H} = 0. However this means that the Eq. 2.16 will be zero, which englobes both the classical energy density plus the effective quantum contribution, which will be in fact relevant at this time. Thus, this condition does not imply that $\rho + p$ = 0 and the model does not suffer from ghost instability problems. We have added the calculation of the second Friedmann equation in section 2.2

2 - d) In Ref 33, it is given an initial wave function by Equation 44. However this is an initial condition set at the bounce. We see below Eq. 45 that the authors afirm: "Note that this solution has no singularities for whatever initial value of $a_b \neq 0$, and
tends to the classical solution when $T \rightarrow \pm T$." The initial condition set at the bounce assures that there are no singularities in the model.

Moreover, our model starts with $T = -\infty$ so there is a contracting phase for the scale factor. We have explicitly defined the initial time in the paragraph below Eq. 2.12.

2 - e) From our definitions, the time variable does not depend on a scalar field, it depends on the fluid's energy density, which is also perturbed.

SANDRO ADICIONAR

3- DEMETRIO ADICIONAR

4 - In Ref.~\cite{vitenti}, it is shown that for a Friedmann background with scalar perturbations, an appropriate gauge choice ensures the perturbative series remains valid in the bounce model. In our study, we focus on this model, leaving the exploration of a homogeneous and anisotropic background, specifically Bianchi-I models, for future work. A brief discussion of this possibility has been included in the conclusion.

5 - In this work, $c$ is the speed of light, $\bar{\rho}$ is the matter energy density and $\delta\rho$ is the perturbation to the matter energy density. The aforementioned quantities were properly defined in the main text, along with other quantities that were used before their definitions.

\bibliographystyle{unsrt}
\bibliography{g}
\end{document}

### 2024-10-09 referee_response.tex

\usepackage{xcolor}
\usepackage{mdframed}

\usepackage[hmarginratio=1:1]{geometry}

\begin{document}
We appreciate the interest in our paper and thank you for the time taken to review it. Regarding the points suggested, here are our thoughts on the matter:

\begin{enumerate}
    \item

    \begin{enumerate}
        \item  To obtain the matter Lagrangian for this model, we employ the Schutz formalism. It assumes that the fluid's velocity can be expressed in terms of potentials of scalar fields, and the fluid's enthalpy is determined by the magnitude of the velocity. This allows us to write the Lagrangian as a functional of these fields. A detailed application of this formalism can be found in Ref.~\cite{fluidgeral}. We have included a brief description of this approach in Section 2.2 and corrected the cited references.

        \item Indeed, if the sound speed were to approach zero, we would encounter a divergence problem. This is why we limit ourselves to small but non-zero equations of state for dust, which is in agreement with the literature. Regarding the $\zeta$ field, we can observe from the initial conditions in equation 3.32, along with the definitions of 
        z and the frequency, that the modes for $\zeta$ will evolve proportionally to 
        $\frac{1}{\sqrt{m \omega_k}} \propto c_s$, meaning that higher-order perturbative terms would evolve similarly.

@@

        In \cite{vitenti}, the authors demonstrate that the first-order terms in the perturbative expansion are indeed smaller than the zero-order terms. To replicate this for higher-order terms, we would need to prove that all third-order terms are smaller than second-order terms in a general manner. Simply comparing one term in the expansion, such as $\zeta^3$ to $\zeta^2$, does not imply that all second-order terms in the Lagrangian are less significant than third-order terms. This kind of proof has not yet been addressed in the literature and is beyond the scope of this paper.

        We have reorganized Sections 4.3 and 4.4 and considered a very small, but non-zero, equation of state for dark matter to analyze the pressureless case.

    \end{enumerate}

    \item

    \begin{enumerate}
        \item Our presentation of the Wheeler-De Witt equation and choice of internal time was too direct and incomplete, as pointed out by the referee. We then added an intermediary step for which one considers the general Wheeler-De Witt equation without a choice of time parameter, so that we only consider $N = a^{3w}$ after identifying a suitable internal time on the Wheeler-De Witt equation.

        \item \textbf{SANDRO ADICIONAR} In our quantum model, we observe that there is a quantum contribution to the usual Friedmann equation. This contribution should not be interpreted as a new fluid but rather as an effective correction. Essentially, we use the classical fluid density with the scale factor following the Bohm trajectory.

@@

        Moreover, if we compute the time in the universe when the quantum contribution reaches 10\% of the classical contribution, we find that this occurs only at scales very close to the bounce. At these timescales, the primordial black holes formed would have such small masses that they would evaporate quickly and thus would not affect our analysis. We restrict our study to times well before the bounce, where such quantum corrections are not significant. Therefore, we only consider $w_{matter}$ and not $w_{eff}$.

        We have added the calculation of the times when the quantum contribution becomes 10\% of the classical one to the paper.

        \item At bouncing point, we have that \bar{H} = 0. However this means that the Eq. 2.16 will be zero, which englobes both the classical energy density plus the effective quantum contribution, which will be in fact relevant at this time. Thus, this condition does not imply that $\rho + p$ = 0 and the model does not suffer from ghost instability problems. We have added the calculation of the second Friedmann equation in section 2.2 .

        \item In Ref 33, it is given an initial wave function by Equation 44. However this is an initial condition set at the bounce. We see below Eq. 45 that the authors afirm: "Note that this solution has no singularities for whatever initial value of $a_b \neq 0$, and
        tends to the classical solution when $T \rightarrow \pm T$." The initial condition set at the bounce assures that there are no singularities in the model.

@@

        Moreover, our model starts with $T = -\infty$ so there is a contracting phase for the scale factor. We have explicitly defined the initial time in the paragraph below Eq. 2.12.

        \item From our definitions, the time variable does not depend on a scalar field, it depends on the fluid's energy density, which is also perturbed. \textbf{SANDRO ADICIONAR}

    \end{enumerate}

    \item As was well observed, our discussion regarding the quantum or classical behavior in 3.3 was not sound. Therefore, we added a comment pointing out that quantum effects are encapsulated on the classical effective scale factor (2.13), which allows one to treat perturbations as quantum and the background as classical in this effective formulation.

    \item In Ref.~\cite{vitenti}, it is shown that for a Friedmann background with scalar perturbations, an appropriate gauge choice ensures the perturbative series remains valid in the bounce model. In our study, we focus on this model, leaving the exploration of a homogeneous and anisotropic background, specifically Bianchi-I models, for future work. A brief discussion of this possibility has been included in the conclusion.

    \item In this work, $c$ is the speed of light, $\bar{\rho}$ is the matter energy density and $\delta\rho$ is the perturbation to the matter energy density. The aforementioned quantities were properly defined in the main text, along with other quantities that were used before their definitions.

\end{enumerate}

### 2024-10-09 main.tex
Added Demetrio's part

\begin{align}
\label{total_hamiltonian_FLRW}
            H & = N\dpar{- \frac{\Pi_{\bar{a}}^{2} }{ 4\bar{a} } + \frac{\Pi_{T}}{\bar{a}^{3w}} } ,
        \end{align}
        where 
        \begin{align}
            \Pi_{\bar{a}} & \equiv \frac{ \partial L}{\partial \dot{\bar{a}}}
            ,\end{align}
        \begin{align}
            \Pi_T & \equiv \frac{ \partial L}{ \partial \dot{T}} 
        ,\end{align}
        are the canonical momenta associated with the scale factor $\bar{a}$ and $T$, 
        the latter being a variable that describes the fluid's evolution. Also, we denote the cosmic time derivative as $\frac{\partial q}{c\partial t} \equiv \dot{q}$, which has a dimension of one over length.
        \textcolor{red}{The momentum $\Pi_T$ is related to the fluid's degree of freedom and is given by $\Pi_T = c^{-1} \tilde{V} \bar{a}^{3(1 + w)} \bar{\rho}$, where $\tilde{V}$ represents the conformal spatial volume, and the fluid's density is a function of enthalpy and entropy.} Remarkably, $\Pi_T$ appears linearly in the Hamiltonian, and its classical equation of motion is given by
        $\dot{T} = { N }/{ \bar{a}^{3w} }$, implying $\dot{T} > 0$ and making $T$ a monotonic function of classical time. Hence, $T$ can be used as a time variable, allowing us to write a Schrödinger-like equation by choosing $N = \bar{a}^{3w} \implies \dot{T} = 1$. For the details, see Ref.~\cite{fluidgeral}.

        Canonical quantization is then performed by promoting classical variables to operators satisfying the canonical commutation relations. This process yields the following Wheeler-DeWitt equation for the wavefunction of the universe  $\Psi(\bar{a}, T)$ \cite{nelson_peter_bouncing_original}
                \begin{equation}
                \label{wdweq}
            i\frac{\partial}{\partial T}\Psi(q,T) = \frac{1}{4}\frac{\partial^{2}}{\partial q^{2}}\Psi(q,T),
        \end{equation}
       such that a specific operator factor ordering was chosen to preserve the symmetries of the classical system \cite{halliwell1990introductory} . Also we introduced the change of variable
        \begin{equation}
            q \equiv \frac{2\bar{a}^{\frac{3}{2}\dpar{1-w} } }{ 3\dpar{1-w } }
        .\end{equation}

Eq.~\eqref{wdweq} resembles a time-reversed free particle Schrödinger equation and, with appropriate boundary conditions, its solutions are wavefunctions of the scale factor $\bar{a}(T)$. We turn to the De Broglie-Bohm interpretation~\cite{mukhanov2005physical,nelson2000bohm}, such that assuming a Gaussian wavefunction $\Psi(\bar{a},T)$, the Bohmian trajectory solution reads
\begin{equation}
\label{bohm_scale_factor}
            \bar{a}(T) = \bar{a}_{B}\dcol{1 + \dpar{ \frac{T}{T_{ B}} }^2 }^{{\frac{1}{3}\frac{1}{\dpar{1-w}} } }  ,
        \end{equation}
        where $\bar{a}_{B}$ is an integration constant that represents the minimum scale factor value and $T_B$ is a small arbitrary constant related to the time scale of the bounce.  \textcolor{red}{Note that our model goes from $T=-\infty$ in the far past untill $T=0$ at the bounce such that we obtain a minimal scale factor at the bounce.} Remarkably, $\bar{a}(t) \neq 0$ for all $t$, which means that this model is non-singular and represents an eternal universe~\cite{nelson_peter_bouncing_original}. For a detailed derivation, see \cite{nelson2021bouncing}.

@@

        Eliminating the time $t$ using \eqref{bohm_time} in \eqref{bohm_hubble}, we find 
        \begin{equation}
        \label{friedman_scale}
            \bar{H}^{2} = \frac{ 4 }{ 9t^{2}_{b}}\dpar
{ \frac{ \bar{a}^{3}_{b} }{ \bar{a}^{3} } - \frac{\bar{a}_{b}^{6}}{ \bar{a}^{6} } }\, ,
        \end{equation}
        which is equivalent to the Friedmann equations
        \begin{equation}\label{effective_friedmann}
            \bar{H}^{2} = \frac{ \kappa c^{2} }{ 3 }\bar\rho - \bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6},
        \end{equation}
        where $\bar\rho$ is the dust energy density. \textcolor{red}{Also, from the derivative of Eq.~\eqref{friedman_scale},
        \begin{align}
         c\dot{\bar{H}} = - \frac{\kappa c^{2}}{2} \bar\rho + 3\bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}    
        .\end{align}
        }
 In the above equations, there is an additional term $- \bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}$ when compared to the usual Friedmann equation for a classical universe. In our model, the scale factor dynamics mirrors a typical Friedmann equation with a total energy density $\rho_{T} = \bar\rho_{d} + \rho_{q}$, where $\rho_{d}$ represents the dust energy density and $\rho_{q} = -\Omega\rho_{c}\bar{a}^{-6}$ denotes an effective negative energy density accounting for quantum effects that circumvent the primordial singularity~\cite{vitenti2012large}. For large-scale factors $\bar{a} \gg \bar{a}_{b} $, Eq.~\eqref{effective_friedmann} and Eq.~\eqref{bohm_scale_factor} reduce to the classical dust dominated flat FLRW universe.
        \textcolor{red}{For instance, let us compute the time for which the quantum contribution is 10\% of the classical one. That means, from Eqs.~\eqref{friedman_scale} and \eqref{bohm_scale_factor} , that $\bar{a} = 0.1 a_b$ and $1 + \left(\frac{T}{T_B}\right)^2 = 1000$ for dust. In our dimensionless units, defined in Eq.~\eqref{tauparameter}, this would result in a time $\tau_a \approx -4.16$, which is extremely close to the bouncing point and it is not relevant for our analysis. Thus, effectively we still have a classical behavior with an energy density following the scale factor from the Bohm trajectory, and the quantum corrections only become relevant when $t \approx t_b$.}
         With this, we conclude the analysis of our background quantum bouncing model and move to its associated perturbations. In the next sections, we will consider the above quantum bouncing model with $w\approx 10^{-10}$ and $x_b = 10^{35}$ and we solve the equations numerically without any approximation.

@@

In this model, the background is characterized by a quantum contracting phase where an effective quantum fluid dominates near the bounce. Consequently, we seek quantized perturbations for consistency, although both theories can be viewed independently. In particular, the use of quantized perturbations evolving on classical backgrounds has been widely used since the results of Mukhanov and Chibisov \cite{mukhanov1981quantum} and Hawking \cite{hawking1982quantum_fluctuations} to derive the power spectrum of primordial perturbations, which are in turn used to describe the formation of structure on our universe.

>>>>

\begin{align}
\label{total_hamiltonian_FLRW}
            {\cal H} & = N\dpar{- \frac{\Pi_{\bar{a}}^{2} }{ 4\bar{a} } + \frac{\Pi_{T}}{\bar{a}^{3w}} } ,
        \end{align}
        where 
        \begin{align}
            \Pi_{\bar{a}} & \equiv \frac{ \partial L}{\partial \dot{\bar{a}}}
            ,\end{align}
        \begin{align}
            \Pi_T & \equiv \frac{ \partial L}{ \partial \dot{T}} 
        ,\end{align}
        are the canonical momenta associated with the scale factor $\bar{a}$ and $T$, 
        the latter being a variable that describes the fluid's evolution. Also, we denote the cosmic time derivative as $\frac{\partial q}{c\partial t} \equiv \dot{q}$, which has a dimension of one over length.
        \textcolor{red}{The momentum $\Pi_T$ is related to the fluid's degree of freedom and is given by $\Pi_T = c^{-1} \tilde{V} \bar{a}^{3(1 + w)} \bar{\rho}$, where $\tilde{V}$ represents the conformal spatial volume, and the fluid's density is a function of enthalpy and entropy.} 
        {\color{red}
        Canonical quantization is then performed by promoting classical variables to operators satisfying the canonical commutation relations\footnote{\color{red} Note that we quantize only the dynamical variables $a,T$ while $N$ is kept classical, for it is not a dynamical degree of freedom.}. This process yields the following Wheeler-DeWitt equation for the wavefunction of the universe  $\Psi(\bar{a}, T)$ \cite{nelson_peter_bouncing_original}
        %
        \begin{equation}
                \label{wdweq}
            N\dcol{ i\frac{\partial}{\partial T}\Psi(q,T) - \frac{1}{4}\frac{\partial^{2}}{\partial q^{2}}\Psi(q,T) } = 0 ,
        \end{equation}
        %
        such that a specific operator factor ordering was chosen to preserve the symmetries of the classical system \cite{halliwell1990introductory} . Also we introduced the change of variable
        \begin{equation}
            q \equiv \frac{2\bar{a}^{\frac{3}{2}\dpar{1-w} } }{ 3\dpar{1-w } }
        .\end{equation}
        %
        Note that the right hand side of side of said equation always  vanishes due to the classical Hamiltonian constraint ${\cal H}=0$. Therefore, if one tries to interpret the right hand it as a usual Schrödinger equation with $\propto \partial \Psi / \partial T$ where $T$ is a time parameter, one sees that the wavefunction does not evolve on time, which is the well known Problem of Time in Quantum Gravity.

        A possible way to evolve the quantum state is to consider an internal time parameter. First, note that the operator $\hat{\Pi}_T$ appears linearly in the Hamiltonian, and the classical equation of motion of its conjugate variable is given by
        $\dot{T} = { N }/{ \bar{a}^{3w} }$, implying $\dot{T} > 0$ and making $T$ a monotonic function of classical time. Hence, $T$ can be used as a time variable, allowing us to write a Schrödinger-like equation by using our freedom to choose the lapse function as $N = \bar{a}^{3w} \implies \dot{T} = 1$. For the details, see Ref.~\cite{fluidgeral}. We then obtain
        %
                \begin{equation}
                \label{wdweq}
            i\frac{\partial}{\partial T}\Psi(q,T) = \frac{1}{4}\frac{\partial^{2}}{\partial q^{2}}\Psi(q,T),
        \end{equation} }

Eq.~\eqref{wdweq} resembles a time-reversed free particle Schrödinger equation and, with appropriate boundary conditions, its solutions are wavefunctions of the scale factor $\bar{a}(T)$. We turn to the De Broglie-Bohm interpretation~\cite{mukhanov2005physical,nelson2000bohm}, such that assuming a Gaussian wavefunction $\Psi(\bar{a},T)$, the Bohmian trajectory solution reads
\begin{equation}
\label{bohm_scale_factor}
            \bar{a}(T) = \bar{a}_{B}\dcol{1 + \dpar{ \frac{T}{T_{ B}} }^2 }^{{\frac{1}{3}\frac{1}{\dpar{1-w}} } }  ,
        \end{equation}
        where $\bar{a}_{B}$ is an integration constant that represents the minimum scale factor value and $T_B$ is a small arbitrary constant related to the time scale of the bounce.  \textcolor{red}{Note that our model goes from $T=-\infty$ in the far past untill $T=0$ at the bounce such that we obtain a minimal scale factor at the bounce.} Remarkably, $\bar{a}(T) \neq 0$ for all $T$, which means that this model is non-singular and represents an eternal universe~\cite{nelson_peter_bouncing_original}. For a detailed derivation, see \cite{nelson2021bouncing}.

@@

        Eliminating the time $t$ using \eqref{bohm_time} in \eqref{bohm_hubble}, we find 
        \begin{equation}
        \label{friedman_scale}
            \bar{H}^{2} = \frac{ 4 }{ 9t^{2}_{b}}\dpar
{ \frac{ \bar{a}^{3}_{b} }{ \bar{a}^{3} } - \frac{\bar{a}_{b}^{6}}{ \bar{a}^{6} } }\, ,
        \end{equation}
        which is equivalent to the Friedmann equations
        \begin{equation}\label{effective_friedmann}
            \bar{H}^{2} = \frac{ \kappa c^{2} }{ 3 }\bar\rho - \bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6},
        \end{equation}
        where $\bar\rho$ is the dust energy density. \textcolor{red}{Also, from the derivative of Eq.~\eqref{friedman_scale},
        \begin{align}
         c\dot{\bar{H}} = - \frac{\kappa c^{2}}{2} \bar\rho + 3\bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}    
        .\end{align}
        }
 In the above equations, there is an additional term $- \bar{H}^{2}_{0}\Omega_{q0}\bar{a}^{-6}$ when compared to the usual Friedmann equation for a classical universe. In our model, the scale factor dynamics mirrors a typical Friedmann equation with a total energy density $\rho_{T} = \bar\rho_{d} + \rho_{q}$, where {\color{red}$\rho_{d}$ represents the dust energy density and $\rho_{q} = -\Omega\rho_{c}\bar{a}^{-6}$ denotes an effective negative energy density accounting for quantum effects that circumvent the primordial singularity}~\cite{vitenti2012large}. For large-scale factors $\bar{a} \gg \bar{a}_{b} $, Eq.~\eqref{effective_friedmann} and Eq.~\eqref{bohm_scale_factor} reduce to the classical dust dominated flat FLRW universe.
        \textcolor{red}{For instance, let us compute the time for which the quantum contribution is 10\% of the classical one. That means, from Eqs.~\eqref{friedman_scale} and \eqref{bohm_scale_factor} , that $\bar{a} = 0.1 a_b$ and $1 + \left(\frac{T}{T_B}\right)^2 = 1000$ for dust. In our dimensionless units, defined in Eq.~\eqref{tauparameter}, this would result in a time $\tau_a \approx -4.16$, which is extremely close to the bouncing point and it is not relevant for our analysis. Thus, effectively we still have a classical behavior with an energy density following the scale factor from the Bohm trajectory, and the quantum corrections only become relevant when $t \approx t_b$.}
         With this, we conclude the analysis of our background quantum bouncing model and move to its associated perturbations. In the next sections, we will consider the above quantum bouncing model with $w\approx 10^{-10}$ and $x_b = 10^{35}$ and we solve the equations numerically without any approximation.

@@

In this model, the background is characterized by a quantum contracting phase where an effective quantum fluid dominates near the bounce. Consequently, we seek quantized perturbations for consistency, although both theories can be viewed independently. {\color{red} Since we are considering the De Broglie-Bohm interpretation for the background, quantum effects are described in terms of the effective classical scale factor  \eqref{bohm_scale_factor}, so that we may treat quantum perturbations evolving on said classical scale factor \cite{Vitenti2013}}. In particular, the use of quantized perturbations evolving on classical backgrounds has been widely used since the results of Mukhanov and Chibisov \cite{mukhanov1981quantum} and Hawking \cite{hawking1982quantum_fluctuations} to derive the power spectrum of primordial perturbations, which are in turn used to describe the formation of structure on our universe.
